---
title: "Final Assignment"
author: "Gonzalo de Quesada, email: gonzalo.dequesadaalzamora@helsinki.fi"
email: "gonzalo.dequesadaalzamora@helsinki.fi"
date: "12 de diciembre de 2017"
output: 
html_document:
    theme: cosmo
    toc: true
    toc_depth: 2
    fig_caption: true
    fig_width: 6
    fig_height: 4
    code_folding: hide
---
## Introduction

## Data

The dataset comes from the United Nations Development Programme. this dataset is the product of two different datasets used to calculate the human development index. for more information about the data you can visit the United Nations Development Programme [website](http://hdr.undp.org/en/content/human-development-index-hdi) and also check the [technical notes](http://hdr.undp.org/sites/default/files/hdr2015_technical_notes.pdf).

Lets inspect the structure of the dataset.it contains 177 observations and 6 variables.

```{r, data, message=FALSE}
library(GGally)
library(dplyr)
library(corrplot)
library(ggplot2)
human <- read.csv("C:/Users/D E L L/Documents/GitHub/IODS-project/Data/human.csv", sep = ",", header = TRUE)
glimpse(human)
```

The variables in the dataset are:

yr.eduexp: Expected years of schooling

lifexp: Life expectancy at birth

GNI: Gross National Income per capita

matermor: Maternal mortality ratio

adobirth: Adolescent birth rate

repreparl: Percetange of female representatives in parliament

We now use the correlation matrix to see the relationship between the variables.

```{r, plot}
ggpairs(human)
cor(human) %>% corrplot
```

We can see a strong positive correlation between life expectancy at birth and 

## Methods

## Results

### Linear Regression

```{r, res}
model <- lm(matermor ~ lifexp + adobirth, data = human)
summary(model)
model
```

```{r, model}
g1 <- ggplot(human, aes(x=matermor, y=lifexp + adobirth))
g1 + geom_point()
```

```{r, plot1}
par(mfrow = c(2,2))
plot(model, which= c(1,2,5))
```

### K-means

```{r, km}
human_scaled <- scale(human)
summary(human_scaled)
dist_eu <- dist(human_scaled)
summary(dist_eu)
```

```{r, kmplot}
k_max <- 10
twcss <- sapply(1:k_max, function(k){kmeans (human_scaled, k)$tot.withinss})
qplot(x = 1:k_max, y = twcss, geom = 'line')
```

```{r, kmplot1}
km <-kmeans(human_scaled, centers = 2)
pairs(human_scaled,col = km$cluster)
```

## Conclusions
